---
title: "Creating new visualizations through joins"
subtitle: "An exploration of Kaggle's Open University Learning Analytics Dataset"
date: "Assigned 1/23/19, Due 1/30/19"
output:
  html_document: 
    toc: true
    toc_float: true
    theme: "journal"
    css: "website-custom.css"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE,
                      echo = FALSE,
                      dev.args = list(bg="transparent"),
                      fig.height = 12)

```

## Background
For this homework we will use data from [kaggle](https://www.kaggle.com). Kaggle is an online data science platform that began as a competitive platform for predictive modeling problems. Private organizations would offer a prize, usually in the range of $25K-$100K, for the individual or team that could provide the best predictive model for their problem. This is still a primary function of the website (see active competitions [here](https://www.kaggle.com/competitions)), but the it has grown and is now also a place to share datasets, compete in open competitions (where you can see others code), and generally learn different facets and approaches to data science in an applied way.

#### Getting Started
You can download the data we'll use for this homework either [from kaggle](https://www.kaggle.com/rocki37/open-university-learning-analytics-dataset) (requires you create an account with kaggle) or [directly from the organization](https://analyse.kmi.open.ac.uk/open_dataset). The latter link is worth examining to learn more about the data, regardless.

A total of seven datasets (tables) related to an online learning analytics project should be included with your download. A number of keys link the various sources of data. I recommend reading through the documentation provided by the organization to better understand these keys.

## Assignment
The primary components of this assignment are highly open-ended. These are to

1. Use at least 3 mutating joins
2. Use at least 1 filtering join
3. Produce at least 3 data displays

Following the completion of these requirements, please reproduce the following figure.

```{r echo = FALSE}
library(tidyverse)
theme_set(theme_minimal() +
	theme(plot.title.position = "plot"))

stu_assess <- read_csv(here::here("kaggle-data", "studentAssessment.csv")) 
stu_info <- read_csv(here::here("kaggle-data", "studentInfo.csv")) 
assessments <- read_csv(here::here("kaggle-data", "assessments.csv")) 

assessments %>%
	count(id_assessment) %>%
	filter(n > 1)

stu_assess %>%
	count(id_student, id_assessment) %>%
	filter(n > 1)

stu_info %>%
	count(id_student, code_module, code_presentation) %>%
	filter(n > 1)

left_join(stu_assess, assessments) %>%
	count(id_student, id_assessment) %>%
	filter(n > 1)

d <- left_join(stu_assess, assessments) %>%
	left_join(stu_info)


# anti join example
d %>%
	count(code_module, id_student)

sub_by_code <- d %>%
	mutate(sub_date_diff = date_submitted - date) %>%
	group_by(id_student, code_module) %>%
	summarize(mean_sub_date_diff = mean(sub_date_diff, na.rm = TRUE))

ggplot(sub_by_code, aes(mean_sub_date_diff)) +
		geom_histogram() +
		facet_wrap(~code_module)

cases_to_remove <- sub_by_code %>%
	filter(mean_sub_date_diff > -10 & mean_sub_date_diff < 10)

sub_by_code %>%
	anti_join(cases_to_remove) %>%
	ggplot(aes(mean_sub_date_diff)) +
		geom_histogram() +
		facet_wrap(~code_module) +
		labs(title = "Submissions outside of 15 days of assessment date")


pd <- d %>%
	filter(assessment_type == "TMA") %>%
	group_by(highest_education, final_result, gender) %>%
	summarize(mean_score = mean(score, na.rm = TRUE)) %>%
	ungroup() %>% 
	mutate(highest_education = fct_reorder(highest_education, mean_score),
	       final_result = fct_reorder(final_result, mean_score)) %>%
	rename(Gender = gender)

ggplot(pd, aes(highest_education, mean_score)) +
	geom_line(color = "gray40", size = 3) +
  geom_point(aes(color = Gender), size = 5) +
	coord_flip() +
	facet_wrap(~final_result, ncol = 1) +
	scale_color_brewer(palette = "Paired") +
	labs(y = "Average Score",
	     x = "Highest Education",
	     title = "Average TMA Scores",
	     subtitle = "Results displayed by Education, gender, and Final Result Designation",
	     caption = "Data from Kuzilek, Hlosta, & Zdrahal (2017). \nhttps://analyse.kmi.open.ac.uk/open_dataset") 
``` 

## Extra credit
For one point extra credit, modify the lines so the plot looks like the below

```{r extracredit}
pd <- pd %>% 
  pivot_wider(
    names_from = Gender,
    values_from = mean_score
  ) %>% 
  mutate(f_higher = ifelse(`F` > M, "Female", "Male")) %>% 
  pivot_longer(
    cols = c(`F`, M), 
    names_to = "Gender",
    values_to = "mean_score"
  ) %>% 
  mutate(Gender = ifelse(Gender == "F", "Female", "Male"))

ggplot(pd, aes(highest_education, mean_score)) +
	geom_line(aes(color = f_higher), size = 3) +
  geom_point(aes(fill = Gender), shape = 21, stroke = 0, size = 5) +
	coord_flip() +
	facet_wrap(~final_result, ncol = 1) +
	scale_color_brewer("Higher Mean", palette = "Pastel1") +
  scale_fill_brewer(palette = "Paired") +
	labs(y = "Average Score",
	     x = "Highest Education",
	     title = "Average TMA Scores",
	     subtitle = "Results displayed by Education, gender, and Final Result Designation",
	     caption = "Data from Kuzilek, Hlosta, & Zdrahal (2017). \nhttps://analyse.kmi.open.ac.uk/open_dataset") +
  theme(legend.position = "bottom",
        legend.title = element_text(size = 11),
        legend.text = element_text(size = 8))
```


