---
title: "Lab 2"
subtitle: "Mapping data to aesthetics"
date: "Assigned 1/16/19, Due 1/21/19"
output:
  html_document: 
    toc: true
    toc_float: true
    theme: "journal"
    css: "website-custom.css"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE,
                      echo = FALSE,
                      fig.width = 10,
                      fig.height = 10,
                      dev.args = list(bg="transparent"))
library(tidyverse)

d <- rio::import("https://github.com/rfordatascience/tidytuesday/raw/master/data/2018/2018-06-26/week13_alcohol_global.csv",
                 setclass = "tbl_df")

theme_set(theme_minimal(base_size = 15))
myblue <- "#409FDF"


td <- d %>%
  pivot_longer(
    cols = contains("_servings"),
    names_to = "type",
    names_pattern = "(.+)_servings",
    values_to = "servings"
  ) %>%
  rename(total_liters = total_litres_of_pure_alcohol)

ggplot(td, aes(servings)) +
  geom_histogram(bins = 10,
                 color = "white",
                 fill = myblue,
                 alpha = 0.7)

ggplot(td, aes(servings)) +
  geom_histogram(bins = 50,
                 color = "white",
                 fill = myblue,
                 alpha = 0.7)

ggplot(td, aes(servings)) +
  geom_histogram(bins = 35,
                 color = "white",
                 fill = myblue,
                 alpha = 0.7)

ggplot(td, aes(servings)) +
  geom_density(bw = 1,
               color = "white",
               fill = myblue,
               alpha = 0.7)

ggplot(td, aes(servings)) +
  geom_density(bw = 5,
               color = "white",
               fill = myblue,
               alpha = 0.7)

ggplot(td, aes(servings)) +
  geom_density(bw = 10,
               color = "white",
               fill = myblue,
               alpha = 0.7)

ggplot(td, aes(servings)) +
  geom_density(color = "white",
               fill = myblue,
               alpha = 0.7)

ggplot(td, aes(servings)) +
  geom_histogram(color = "white",
                 bins = 35) +
  facet_wrap(~type)

ggplot(td, aes(servings)) +
  geom_density(aes(fill = type),
               color = "white",
               alpha = 0.25) +
  scale_fill_brewer(palette = "Set2")


ggplot(td, aes(servings, type)) +
  ggridges::geom_density_ridges(fill = myblue,
                                color = "white")

ggplot(td, aes(servings, type)) +
  ggridges::geom_density_ridges(fill = myblue,
                                color = "white") +
  xlim(0, 500)

ggplot(td, aes(type, servings)) +
  ggforce::geom_sina()
```

# Overview
The purpose of this lab is to get you to practice more with ggplot, 
specifically with visualizing distributions and amounts.

## Data
We'll work with old [#tidytuesday](https://twitter.com/search?q=%23tidytuesday&src=tyah) data again, this time looking at global
alcohol consumption. The data were originally published by 
[fivethirtyeight](https://github.com/fivethirtyeight/data) to accompany their 
[story](https://fivethirtyeight.com/features/dear-mona-followup-where-do-people-drink-the-most-beer-wine-and-spirits/) 
on the topic.

The data can be read directly into R from the web using  `rio::import` with the following link: https://github.com/fivethirtyeight/data/raw/master/alcohol-consumption/drinks.csv

# Assignment
Start a new R Markdown document - you can choose to house this lab on 
   GitHub or not. Either way is fine.

1. Write the code to load the data in directly from the web, using the link 
   above.
1. Explore the overall distribution of total alcohol consumed. 
  + Histogram 
    - play with the number of bins 
    - settle on a number of bins and provide a brief justification in text
  + Density plot
    - compare the default bandwidth with a few other selections entered 
      manually.
    - settle on a bandwidth selection, and justify why briefly in text  

Reorganize the data so you have a `type` and `servings` column. 

1. Evaluate the distribution of servings by type using at least three 
   different methods. Provide a brief substantive conclusion.
1. Reproduce the following plot. It does not have to be exact, but should be close

```{r echo = FALSE}
top25 <- td %>%
  mutate(type = stringr::str_to_title(type)) %>%
  arrange(desc(total_liters)) %>%
  slice(1:(25*3))


pd <- top25 %>%
  group_by(country) %>%
  mutate(tot_servings = sum(servings)) %>%
  ungroup() %>%
  mutate(servings_prop = servings/tot_servings)

ggplot(pd, aes(country, servings_prop)) +
  geom_col(aes(fill = type),
           alpha = 0.8) +
  coord_flip() +
  scale_fill_brewer("Alcohol Type", 
                    palette = "Paired",
                    breaks = c("Wine", "Spirit", "Beer")) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = "",
       y = "Proportion of total alcohol consumed",
       title = "Alcohol consumption by type",
       subtitle = "Top 25 countries displayed (according to total liters consumed)") +
  theme(legend.position = c(0.8, -0.1),
        legend.direction = "horizontal",
        plot.title.position = "plot",
        plot.margin = margin(b = 2, r = 2, unit = "cm"))
```

### Extra Credit
For up to one point extra credit, reorder countries according to wine consumption, so the plot looks as follows

```{r echo = FALSE}
wine <- filter(pd, type == "Wine") %>%
  mutate(country = forcats::fct_reorder(country, servings))

pd %>%
  mutate(country = factor(country, levels = levels(wine$country))) %>%
ggplot(aes(country, servings_prop)) +
  geom_col(aes(fill = type), 
           alpha = 0.8) +
  coord_flip() +
  scale_fill_brewer("Alcohol Type", 
                    palette = "Paired",
                    breaks = c("Wine", "Spirit", "Beer")) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = "",
       y = "Proportion of total alcohol consumed",
       title = "Alcohol consumption by type",
       subtitle = "Top 25 countries displayed (according to total liters consumed)") +
  theme(legend.position = c(0.8, -0.1),
        legend.direction = "horizontal",
        plot.title.position = "plot",
        plot.margin = margin(b = 2, r = 2, unit = "cm"))

```

# Finishing up
When you have finished the above, upload your rendered (knit) HTML file to canvas.